Towards a standardized evaluation of imputation methodology

### 2022-03-31 Formatted/Formatted.tex
Apply suggested changes (feedback Gerko 30-03), save result as LastDiscussed.tex for latexdiff

The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been. A popular variant is multiple imputation \citep{rubi76}, whereby each missing value is imputed multiple times. The completed data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference \citep[using `Rubin's rules'][]{rubi87}. In contrast to ad hoc methods for dealing with missing values (e.g. list-wise deletion, mean imputation, regression imputation, last observation carried forward, indicator method), multiple imputation properly takes into account the sources of uncertainty that are related to the missingness problem.

@@

As of today, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods. Typically, the developer of a new imputation routine does some tests by simulations, but these tests differ across developers. This brings forth a chain of potential problems in the objective assessment of imputation method performance, within and across studies, which may lead to sub-optimal use of imputation in practice. To demonstrate the broad impact of these problems, we subdivide the problems into the following three distinct categories: problems with data generation, problems with missingness generation, and problems with performance evaluation. We further detail the impact each of these problems may have on the validity of our evaluations.

@@

Evaluating imputation methodology requires a missing data problem to be solved. After establishing a comparative truth from a data generating mechanism, some form of missingness therefore has to be induced. Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like
\begin{quote}
\textit{We generated missing data following a missing at random missingness mechanism.}
\end{quote}
This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness mechanisms and missingness patterns. Missingness mechanisms describe the relationship between missing entries and observed data values \citep[generally categorized into MCAR, MAR and MNAR mechanisms c.f.][see Table \ref{table:mech}]{rubi76}.  Missingness patterns concern the location of missing entries across incomplete data \citep[][p. 8]{litt20}. Under this definition, there is a row-wise element to the missingness pattern describing which variables are jointly observed, and a column-wise element encompassing the amount of missingness in the data.

@@

Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables. It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways. Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to sub-optimal advice in practice. For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that list-wise deletion outperforms their imputation method(s). Such a conclusion would, unfortunately, only translate to data that adhere to the same special case \citep[described in][\S 2.7]{fimd}, while biasing inferences in other cases. Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e. a pattern with uniformly increasing missingness proportions along variables;][]{litt20}. If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations. A final example is multivariate missingness that is generated using step-wise univariate missingness induction. The resulting incomplete data may then not have the intended or expected statistical properties \citep[e.g. skewness;][]{ampute}. This complicates any definitive conclusions about imputation method performance in relation to the data generating mechanism.

@@

\begin{table}[tb]
\begin{center}
\caption{Missingness mechanisms. [Add PMAR?]}
\label{table:mech}
\begin{tabular}{lll}
\hline
Mech. & Interpretation and consequences of the mechanism \\
\hline  
MCAR  & Missing completely at random: The probability to be missing is the same for all \\
      & cases. In other words, the missing values are missing at random and the observed \\
      & values are observed at random. \\
MAR   & Missing at random: The probability to be missing is the same within groups of \\
      & cases defined by the observed data only. The essence is that observed relations in \\
      & the data inform the missingness, such that MCAR may be assumed within the \\ 
      & observed groups. \\
MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
      & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
      & inferred from the observed data. This is considered non-ignorable missingness \\
      & \citep[see e.g.][]{rubi76}. \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

Although MAR missingness is often considered as simulation condition, the problem of spurious MAR is generally overlooked. With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g. weight is made incomplete based on observed gender to emulate a situation wherein one gender is less likely to disclose their weight). These relations in the observed data may, however, be weak or non-existent. If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid. The most extreme example would be when MAR is induced from data without multivariate relations. The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random. In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p. 48]{fimd}: the missingness does not depend on the incomplete variable. This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.

Inducing a MAR mechanism comes with another issue, namely the type or functional form of the missingness model. Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model \citep{scho18}. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

@@

@@

The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models (i.e. the complete data model used to estimate a target or estimand). When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}. Congeniality is difficult to diagnose, but ignoring it may yield sub-optimal imputations. Methods to asses the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}, which is an infeasible endeavor in many simulation studies. % [All imputation models should be congenial, but this is difficult to test and often ignored. Checks of imputation model fit are often omitted. Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others. Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies. Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]  
Therefore, the performance of imputation procedures on distributional properties is often ignored in simulation studies. Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers. For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large. While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.

@@

If the goal is inference--not prediction--then the uncertainty about estimates needs to be quantified. A single imputation would then not suffice. The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied. After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p. 591]{neym34}. Merely focusing on distance measures would open the door for invalid inferences.

Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification. For many outcome-driven imputation task a single valid imputation may suffice. Measures of predictive accuracy or the RMSE (root mean squared error) of predicted values could then serve as sensible evaluation criteria.

@@

Standardizing the evaluation of imputation methodology requires simulators--like ourselves--to consider a couple of key aspects in their simulation workflows. Table \ref{table:steps} outlines some suggested steps to adopt. Please note that there is--of course--no `silver bullet' and that we do not claim to present a universally applicable approach. Our suggestions are the accumulated result of scientific literature, research experience, and discussion. For an excellent overview of general best practices for method evaluation by means of simulation, see \citet{morr18}. We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology.

@@

Before setting up a simulation study, the simulator should clearly define the scope of their evaluations. A simulation study aimed at imputation model comparison will typically have a  different design than one aimed at establishing the inferential validity of a single (novel) method. The choice of imputation method(s) under evaluation is naturally intertwined with the simulation study's aims and set-up.

@@

Some general choices in the simulation set-up to consider at this stage are the simulation design and the number of simulation repetitions. The simulation design should ideally be fully factorial, i.e. varying each simulation condition against all other conditions. There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions). If there are several imputation method under evaluation, the simulation design should accommodate that each method is applied to each incomplete data set. This is in contrast to a design with as many incomplete data sets as imputation methods in every simulation repetition. Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation which makes for fairer comparisons. The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g. as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.

@@

The simulator should choose their data generating mechanism(s) in line with the study scope. Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived. A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest. The benefit of design-based simulation is the ability to use real-life observed data structures.

Estimand(s) or other simulation targets should defined in the context of the study aim and data-generating mechanism. [As a general rule, it would be wise to include one or more multivariate estimands. It is a realistic expectation for imputation methods to yield valid univariate estimates but also to capture multivariable relations in the data. Imputation methods that fail to do so, should not be considered general purpose methods (e.g., mean imputation).]

@@

First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions). This mechanism may be used as a reference condition. Valid imputations under MCAR is a minimal requirement for any [multipurpose] imputation method. We also know that, under MCAR, the distribution of imputed values should be equivalent to the distribution of observed values for a given variable, which offers an indication of imputation model (mis)fit. 
% First, one should always consider MCAR missingness [as reference, because all imputation methods should yield valid inference under MCAR and we can use this mechanism to check imputation model fit] %, i.e. the scenario where the missing values are missing at random and the observed values are observed at random. % Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.

Next, missing data should be induced conform a model that is dependent on the observed data (i.e. a MAR mechanism). Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully. A straightforward technique for inducing univariate MAR missingness is described in \citet[][p. 63]{fimd}, generalizations to multivariate MAR missingness can be found in \citet{ampute}, and if the missingness is to be induced in longitudinal data autoregressive MAR models \citep[e.g. cf.][model 2 and model 3]{shara2015randomly} can be useful. It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness. The effects of different types of MAR mechanisms are described in \citet{scho18}.% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook results due to the focus on a single MAR mechanism.

Third, a non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice. Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time. It would, therefore, be wise to include MNAR missingness in the simulation study just in case, or to perform sensitivity analyses to get an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.

Apart from the missingness mechanisms, the missingness pattern must be varied--in particular the missingness proportion. Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process. Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data. The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness. If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption. Alternatively, the validity of assumptions become increasingly important when the missingness increases. Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions. We, therefore, propose to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness (see Table \ref{table:prop}). % Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.

@@

\subsection{Apply missing data method(s) and analysis model(s)}

@@

The imputation method(s) under evaluation should be applied to the generated incomplete data. It is important to choose the imputation model parameters carefully (e.g., number of imputations, number of iterations, regularization, etc.). After imputation, an analysis model may be fitted to estimate the estimand(s).

%[Use CCA as benchmark method: the imputation method should outperform CCA. Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matrix, regularization if applicable, etc.). Estimate estimand(s) using the analysis model, use correct pooling rules.]

Apart from the imputation method(s), list-wise deletion should be considered as missing data method too. We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating multiple imputation performance. List-wise deletion may therefore serve as a benchmark method: out-performing it should be a minimal requirement for the imputation method(s). %it is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.

@@

% = evaluation of imputation generating process
The absence of non-convergence in the imputation models is a minimal requirement for imputation methods. Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}. The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.

@@

The fit of the imputation model may be verified with the help of a posterior predictive check \citep[PPC][]{nguy17, zhao22}. A PPC in the context imputation methodology could be the over-imputation of observed data values. The reasoning is that if the over-imputed values have  equivalent statistical properties to the true data values, the imputation model fits the observed part of the incomplete data well. We might then assume that the imputation model would be able to produce good imputations for missing values as well.

@@

[Some suggested minimal requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis (i.e. performance after list-wise deletion).]

@@

It could be convincing to demonstrate a method's applicability to real-world missing data problems. This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set. Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data. However, the real-world missingness would then be omitted.

>>>>

The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been. A popular variant is multiple imputation \citep{rubi76}, whereby each missing value is imputed multiple times. The completed data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference using Rubin's rules \citep[][p. 76]{rubi87}. In contrast to ad hoc methods for dealing with missing values (e.g. list-wise deletion, mean imputation, regression imputation, last observation carried forward, indicator method), multiple imputation properly takes into account the sources of uncertainty that are related to the missingness problem.

@@

As of today, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods. Typically, the developer of a new imputation routine does some tests by simulations, but these tests differ across developers. This brings forth a chain of potential problems in the objective assessment of imputation method performance, within and across studies, which may lead to sub-optimal use of imputation in practice. To demonstrate the broad impact of these problems, we subdivide the problems into the following three distinct categories: problems with data generation, problems with missingness generation, and problems with performance evaluation. We further detail the impact each of these problems may have on the validity of evaluations.

@@

Evaluating imputation methodology requires a missing data problem to be solved. After establishing a comparative truth from a data generating mechanism, some form of missingness therefore has to be induced. Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like:
\begin{quote}
\textit{We generated missing data following a missing at random missingness mechanism.}
\end{quote}
This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness mechanisms and missingness patterns. Missingness mechanisms describe the relationship between missing entries and observed data values and are generally categorized into MCAR, MAR and MNAR mechanisms \citep[][see Table \ref{table:mech}]{rubi76}. [TODO: maybe remove mechanisms here?] Missingness patterns concern the location of missing entries across incomplete data \citep[][p. 8]{litt20}. Under this definition, there is a row-wise element to the missingness pattern describing which variables are jointly observed, and a column-wise element encompassing the amount of missingness in the data.

@@

Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables. It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways. Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to sub-optimal advice in practice. For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that list-wise deletion outperforms their imputation method(s). Such a conclusion would, unfortunately, only translate to data that adhere to the same special case \citep[described in][\S 2.7]{fimd}, while biasing inferences in other cases. Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e. a pattern with uniformly increasing missingness proportions along variables;][]{litt20}. If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations. A final example is multivariate missingness that is generated using step-wise univariate missingness induction. The resulting incomplete data may then not have the desired statistical properties \citep[e.g. skewness;][]{ampute}. This complicates any definitive conclusions about imputation method performance in relation to the data generating mechanism.

@@

\begin{table}[tb]
\begin{center}
\caption{Missingness mechanisms. [TODO: add PMAR?]}
\label{table:mech}
\begin{tabular}{lll}
\hline
Mech. & Interpretation and consequences of the mechanism \\
\hline  
MCAR  & Missing completely at random: The probability to be missing is the same for all \\
      & cases. In other words, the missing values are missing at random and the observed \\
      & values are observed at random. \\
MAR   & Missing at random: The probability to be missing is the same within groups of \\
      & cases defined by the observed data only. The essence is that observed relations in \\
      & the data inform the missingness, such that MCAR may be assumed within the \\ 
      & observed groups. \\
MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
      & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
      & inferred from the observed data. This is considered non-ignorable missingness \\
      & \citep[see e.g.][]{rubi76}. \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

Although MAR missingness is often considered as simulation condition, the problem of spurious MAR is generally overlooked. With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g. weight is made incomplete based on observed gender to emulate a situation wherein one gender is less likely to disclose their weight). These relations in the observed data may, however, be weak or non-existent. If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid. The most extreme example would be when MAR is induced from data without multivariate relations. The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random. In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p. 48]{fimd}: the missingness does not depend on the incomplete variable. This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is pointless.

Inducing a MAR mechanism presents the simulator with another choice, namely the type or functional form of the missingness model. Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model \citep{scho18}. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

@@

The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models (i.e. the complete data model used to estimate a target or estimand). When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}. Congeniality is difficult to diagnose, but ignoring it may yield sub-optimal imputations. Methods to asses the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}, which is an infeasible endeavor in many simulation studies. Therefore, the performance of imputation procedures on distributional properties is often ignored in simulation studies. Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers. For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large. While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.

If the goal is inference--not prediction--then the uncertainty about estimates needs to be properly quantified. A single imputation would then not suffice. The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied. After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p. 591]{neym34}. Merely focusing on the discrepancy between the true and imputed data would open the door for invalid inferences.

Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification. For many outcome-driven imputation task a single valid imputation may perhaps suffice. We would like to note, however, that using measures of predictive accuracy or the RMSE (root mean squared error) of predicted values as the evaluation criterion may increase the rate of false positives \citep[][\S 2.6]{fimd}.

@@

@@

@@

Standardizing the evaluation of imputation methodology requires simulators to consider a couple of key aspects in their simulation workflows. Table \ref{table:steps} outlines some suggested steps to adopt. Please note that there is--of course--no `silver bullet' and that we do not claim to present a universally applicable approach. Our suggestions are the accumulated result of scientific literature, research experience, and discussion. For an excellent overview of general best practices for method evaluation by means of simulation, see \citet{morr18}. We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology.

@@

Before setting up a simulation study, the simulator should clearly define the scope of their evaluations. A simulation study aimed at imputation method comparison will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method. The choice of imputation method(s) under evaluation is naturally intertwined with the simulation aims and set-up.

@@

Some general choices in the simulation set-up to consider at this stage are the simulation design and the number of simulation repetitions. The simulation design should ideally be fully factorial, i.e. varying each simulation condition against all other conditions. There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions). If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set. Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation, which makes for fairer comparisons. The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g. as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.

@@

The simulator should choose their data generating mechanism(s) in line with the study scope. Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived. A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.

Estimand(s) or other simulation targets should be defined in the context of the study aim and data-generating mechanism. As a general rule, it would be wise to include one or more multivariate estimands. It is a realistic expectation for imputation methods to preserve both marginal and conditional distributions with respect to the comparative truth. Imputation methods that fail to do so, should not be considered general purpose methods (e.g., mean imputation).

@@

First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions). This mechanism should be used as a reference condition. Every imputation method should yield valid imputations under MCAR. We also know that, under MCAR, the distribution of imputed values should be equivalent to the distribution of observed values for a given variable, which may offer an indication of imputation model (mis)fit. [TODO: leg uit!]
% First, one should always consider MCAR missingness [as reference, because all imputation methods should yield valid inference under MCAR and we can use this mechanism to check imputation model fit] %, i.e. the scenario where the missing values are missing at random and the observed values are observed at random. % Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.

Next, missing data should be induced conform a model that is dependent on the observed data (i.e. a MAR mechanism). Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully. A straightforward technique for inducing univariate MAR missingness is described in \citet[][p. 63]{fimd}, generalizations to multivariate MAR missingness can be found in \citet{ampute}. If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly} can be useful. It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness. The effects of different types of MAR mechanisms are described in \citet{scho18}.% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook results due to the focus on a single MAR mechanism.

Third, a non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice. Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time. It would, therefore, be wise to include MNAR missingness in the simulation study just in case. Alternatively, sensitivity analyses may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.

Apart from the missingness mechanisms, the missingness pattern must be varied. [TODO: beter bruggetje?] Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process. Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data. The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness. If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption. Alternatively, the validity of assumptions become increasingly important when the missingness increases. Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions. We, therefore, propose to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness (see Table \ref{table:prop}). % Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.

@@

\subsection{Apply methods}

@@

The imputation methods under evaluation should be applied to the generated incomplete data, after which an analysis model may be fitted to estimate the estimand(s). % It is important to choose the imputation model parameters carefully (e.g., number of imputations, number of iterations, regularization, etc.). After imputation, an analysis model may be fitted to estimate the estimand(s).
%[Use CCA as benchmark method: the imputation method should outperform CCA. Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matrix, regularization if applicable, etc.). Estimate estimand(s) using the analysis model, use correct pooling rules.]
In addition, the simulator should always perform complete case analysis. We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance. Complete case analysis may therefore serve as a benchmark method against which performance should be evaluated. Moreover, pairing complete case analysis with MCAR and MAR mechanisms allows the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.% out-performing it should be a minimal requirement for the imputation method(s). %it is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.

@@

% = evaluation of imputation generating process
The absence of non-convergence in the imputation models is a minimum requirement for imputation methods. Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}. The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations might be invalid.

@@

The fit of the imputation model may be verified with the help of a posterior predictive check \citep[PPC][]{nguy17, zhao22}. A PPC in the context of imputation methodology could be the multiple over-imputation of observed data values. The reasoning is that if the over-imputed values have equivalent statistical properties to the true data values, the imputation model fits the observed part of the incomplete data well. We might then assume that the imputation model would be able to produce good imputations for missing values too.

@@

[Some suggested minimum requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis.]

@@

It could be convincing to demonstrate a method's applicability to real-world missing data problems. This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set. Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data. However, the real-world missingness would then be omitted. [TODO: check waar dit al eerder staat!]
